---
title: "BNPdensity_DPprior"
author: "Bystrova Daria"
date: "6/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("BNPdensity")
library(BNPdensity)
library(ggplot2)
```

## Dirichlet prior for Gaussian Mixture

Here we reporduce the example from the (DeBlasi 2015) paper.
The initial data comes from the distribution : uniform mixture of two Gaussians with parameters 
$N(2,0.2)$ and $N(10,0.2)$. Number of samples $N =50$ in the initial example, and the parameter $\alpha =19.6$, such that $\mathbb{E}[K_n] =25$.

```{r mix data}

gen_data<- function(mu1=2,mu2=10, sigma1=0.2, sigma2=0.2, ns=50, prob_vec=c(0.5,0.5)){
  components <- sample(1:2,prob=prob_vec,size=ns,replace=TRUE)
  mus <- c(mu1,mu2)
  sds <- sqrt(c(sigma1,sigma2))
  samples <- rnorm(n=ns,mean=mus[components],sd=sds[components])
  return(samples)
}

sample<- gen_data(ns=50)
#hist(sample,probability=TRUE,breaks=40,col=grey(.9),ylim=c(0,ymax))

```

## Density estimation

Firstly, fit the initial example from the paper:

Number of samples $N =50$
$\alpha =19.6$



```{r BNPmix1 50}
NS=50
DP.alpha=19.6
sample50<- gen_data(ns=NS)
it<-1500
burn<-floor(it*0.3)
result <- MixNRMI2(sample50, probs = c(0.025, 0.5, 0.975) ,Alpha=DP.alpha, Beta=1, Gama=0, distr.k = 1,distr.py0=1, Nit=it, Pbi=0.3)

n_it<-it-burn
df_res <- data.frame(matrix(NA, nrow =NS, ncol =1))
df_res$sample<-sample50

# Plotting density estimate + 95% credible interval
m <- ncol(result$qx)
ymax <- max(result$qx[,m])
par(mfrow=c(1,1))
hist(sample50,probability=TRUE,breaks=40,col=grey(.9),ylim=c(0,ymax),main=paste0("Density estimate for K=2,N=",NS,", alpha =",DP.alpha))
lines(result$xx,result$qx[,1],lwd=2)
lines(result$xx,result$qx[,2],lty=3,col=4)
lines(result$xx,result$qx[,m],lty=3,col=4)
#title(paste0("Density estimate for K=2,N=",50,"alpha =",19.6))

# Plotting number of clusters
#par(mfrow=c(2,1))
plot(1:n_it, result$R,type="l",main="Trace of R")


n_it<-it-burn
df_r <- data.frame(matrix(NA, nrow =n_it, ncol =1))
df_r$clust<-result$R
df_r$it<-1:n_it

 p_est<- ggplot(df_r, aes(x=clust)) +  geom_histogram(bins=10,aes(y=..density..),colour="black")+
   geom_density(aes(y=..density..),color="red",adjust = 2)+labs(title=paste0("Posterior distribution for number of clusters, N=",NS,",CPO=",round(result$cpo,2),",alpha=",DP.alpha)) +
   theme_bw() + theme(axis.text.x = element_text(angle = 0, hjust = 1,size = 10), strip.text = element_text(size = 15),legend.position = "top", plot.title = element_text(hjust = 0.5))
 p_est


```


Result is similar to the paper result!. 

What happens if we take number of sample bigger then 50?
$N=150$




```{r BNPmix1 150}

NS=150
DP.alpha=19.6
sample150<- gen_data(ns=NS)
it<-1500
burn<-floor(it*0.3)
result <- MixNRMI2(sample50, probs = c(0.025, 0.5, 0.975) ,Alpha=DP.alpha, Beta=1, Gama=0, distr.k = 1,distr.py0=1, Nit=it, Pbi=0.3)

n_it<-it-burn
df_res <- data.frame(matrix(NA, nrow =NS, ncol =1))
df_res$sample<-sample50

# Plotting density estimate + 95% credible interval
m <- ncol(result$qx)
ymax <- max(result$qx[,m])
par(mfrow=c(1,1))
hist(sample50,probability=TRUE,breaks=40,col=grey(.9),ylim=c(0,ymax),main=paste0("Density estimate for K=2,N=",NS,", alpha =",DP.alpha))
lines(result$xx,result$qx[,1],lwd=2)
lines(result$xx,result$qx[,2],lty=3,col=4)
lines(result$xx,result$qx[,m],lty=3,col=4)
#title(paste0("Density estimate for K=2,N=",50,"alpha =",19.6))

# Plotting number of clusters
#par(mfrow=c(2,1))
plot(1:n_it, result$R,type="l",main="Trace of R")


n_it<-it-burn
df_r <- data.frame(matrix(NA, nrow =n_it, ncol =1))
df_r$clust<-result$R
df_r$it<-1:n_it

 p_est<- ggplot(df_r, aes(x=clust)) +  geom_histogram(bins=10,aes(y=..density..),colour="black")+
   geom_density(aes(y=..density..),color="red",adjust = 2)+labs(title=paste0("Posterior distribution for number of clusters, N=",NS,",CPO=",round(result$cpo,2),",alpha=",DP.alpha)) +
   theme_bw() + theme(axis.text.x = element_text(angle = 0, hjust = 1,size = 10), strip.text = element_text(size = 15),legend.position = "top", plot.title = element_text(hjust = 0.5))
 p_est


```
 
 
 
 
```{r BNPmix1 150 2}

NS=150
DP.alpha=2
sample150<- gen_data(ns=NS)
it<-1500
burn<-floor(it*0.3)
result <- MixNRMI2(sample50, probs = c(0.025, 0.5, 0.975) ,Alpha=DP.alpha, Beta=1, Gama=0, distr.k = 1,distr.py0=1, Nit=it, Pbi=0.3)

n_it<-it-burn
df_res <- data.frame(matrix(NA, nrow =NS, ncol =1))
df_res$sample<-sample50

# Plotting density estimate + 95% credible interval
m <- ncol(result$qx)
ymax <- max(result$qx[,m])
par(mfrow=c(1,1))
hist(sample50,probability=TRUE,breaks=40,col=grey(.9),ylim=c(0,ymax),main=paste0("Density estimate for K=2,N=",NS,", alpha =",DP.alpha))
lines(result$xx,result$qx[,1],lwd=2)
lines(result$xx,result$qx[,2],lty=3,col=4)
lines(result$xx,result$qx[,m],lty=3,col=4)
#title(paste0("Density estimate for K=2,N=",50,"alpha =",19.6))

# Plotting number of clusters
#par(mfrow=c(2,1))
plot(1:n_it, result$R,type="l",main="Trace of R")


n_it<-it-burn
df_r <- data.frame(matrix(NA, nrow =n_it, ncol =1))
df_r$clust<-result$R
df_r$it<-1:n_it

 p_est<- ggplot(df_r, aes(x=clust)) +  geom_histogram(bins=20,aes(y=..density..),colour="black")+
   geom_density(aes(y=..density..),color="red",adjust = 3)+labs(title=paste0("Posterior distribution for number of clusters, N=",NS,",CPO=",round(result$cpo,2),",alpha=",DP.alpha)) +
   theme_bw() + theme(axis.text.x = element_text(angle = 0, hjust = 1,size = 10), strip.text = element_text(size = 15),legend.position = "top", plot.title = element_text(hjust = 0.5))
 p_est


```
